Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features

### 2016-12-20 paper/structureScene.tex
Edit scattering section

\subsection{Invariance and stability in audio signals}
The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
Indeed, recordings may be shifted locally without affecting their original class memberships.
To discard this superfluous source of variability, signals can be mapped to a time-shifting invariant feature space before training the classifier, eliminating the need for this classifier to explicitly learn this invariance and increasing the robustness of its predictions.

Formally, given a signal $\boldsymbol{x}(t)$, we would like its translation $\boldsymbol{x_c}(t) = \boldsymbol{x}(t-c)$ to be mapped to the same feature vector provided that $|c| \ll T$ for some maximum duration $T$ that specifies the extent of the time-shifting invariance. We can also define more complicated transformations by letting $c$ vary with $t$. In this case, we have $\boldsymbol{x_\tau}(t) = \boldsymbol{x}(t-\tau(t))$ for some function $\tau$, which performs a time-warping of $\boldsymbol{x}(t)$ to obtain $\boldsymbol{x_\tau}(t)$. Time-warpings model various changes, such as small variations in pitch, reverberation, and rhythmic organization of events. These make up an important part of intra-class variability among natural sounds, so a representation that has some insensitivity to such transformations would be very advantageous.

@@

To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\phi(t)$ with cut-off frequency $1/T$ (and so a duration of approximately $T$), to get
\begin{equation}
\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \phi(t),
\end{equation}
which is known as the set of first-order scattering coefficients. They capture the average spectral envelope of $\boldsymbol{x}$ over scales of duration $T$, with the spectral resolution varying with constant Q, as described above. In this way, they are very closely related to the mel-frequency spectrogram and related features, such as MFCCs.

\subsection{Extracting modulations with second-order scattering}
In auditory scenes, amplitude modulations may be caused by a broad variety of rapid mechanical interactions, including collision, friction, turbulent flow, and so on.
At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a local time-shifting invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform followed by a pointwise modulus operator.

@@

The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$. While higher-order scattering coefficients can be calculated, for the purposes of our current, work, the first and second order are sufficient.

>>>>

\subsection{Invariance and stability in audio signals}
The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
Indeed, recordings may be shifted locally without affecting their original class memberships.
To discard this superfluous source of variability, signals are mapped into a time-shifting invariant feature space before training the classifier, eliminating the need for this classifier to explicitly learn this invariance and increasing the robustness of its predictions.

Formally, given a signal $\boldsymbol{x}(t)$, we would like its translation $\boldsymbol{x_c}(t) = \boldsymbol{x}(t-c)$ to be mapped to the same feature vector provided that $|c| \ll T$ for some maximum duration $T$ that specifies the extent of the time-shifting invariance. We can also define more complicated transformations by letting $c$ vary with $t$. In this case, we have $\boldsymbol{x_\tau}(t) = \boldsymbol{x}(t-\tau(t))$ for some function $\tau$, which performs a time-warping of $\boldsymbol{x}(t)$ to obtain $\boldsymbol{x_\tau}(t)$. Time-warpings model various changes, such as small variations in pitch, reverberation, and rhythmic organization of events. These make up an important part of intra-class variability among natural sounds, so it is important for representation to be robust to such transformations.

@@

To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\phi(t)$ with cut-off frequency $1/T$ (and approximate duration $T$), to get
\begin{equation}
\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \phi(t),
\end{equation}
which is known as the set of first-order scattering coefficients. They capture the average spectral envelope of $\boldsymbol{x}$ over scales of duration $T$, with spectral resolution varying with constant Q. In this way, they are closely related to the mel-frequency spectrogram and related features, such as MFCCs.

\subsection{Extracting modulations with second-order scattering}
In auditory scenes, short-time amplitude modulations may be caused by a variety of rapid mechanical interactions, including collision, friction, turbulent flow, and so on.
At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a local time-shifting invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform followed by a pointwise modulus operator.

@@

The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$. While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.
